Election 2024

### 2024-10-30 paper/paper.qmd
reformat the paper so that it looks better

The data used in this paper was gathered from FiveThirtyEight [@FiveThirtyEight] a website that aims to use "data and evidence to advance public knowledge". The programming language for data analysis, visualization and statistical investigation was Python [@Python] along with the packages...

@@

# 2.1 The Dataset

A swing state in the context of a US election is defined as "a state where the number of Democratic and Republican voters is about the same, that has an important influence on the result of the election of the United States President". States that do not fall under his definition are not likely to "swing" to another party as we approach the election, and thus their outcome can be reasonably assumed. Thus the election result is normally decided by which party wins just a handful of states in the electoral college system. By concentrating on these swing states, we aim to capture the most critical and uncertain areas of the electoral map, where voter preferences are most likely to sway the final result. Thus our analysis restricts the datasets to state level polls in the following states: Arizona, Pennsylvania, North Carolina, Georgia, Nevada, Michigan, and Wisconsin [@nyt_swing, @538]. In figure 3 we see the ammount that each party hold according to those sources, with the remaining swing states, seen in the middle of this graph set to determine if either Harris or Trump recieve over 270th electoral votes.

@@

 # 2.2 Variables of Interest
In our logistic regression model predicting the 2024 U.S. Presidential Election outcome, the primary variables of interest are derived from aggregated state-level polling data. The percentage of respondant favouring a candidate become the key variables for prediction. Within the collection of polls in our sample the independent variable `trump_pct` represents the percentage of respondents who indicate support for Donald Trump , and `harris_pct` respectively.
Polls ar taking towards the build up of an election and thus ar taken by repondants at different times. Graph  x shows how the sample of polls are distributed over time in which it is clear that we have a collection of polls from a wide range of periods. Because sentiment changes over time we will adjust the weight that we place on polls that are closer to the election as these will be more indicative of how respondants will be voting on the day.
```{python}
#| echo: false

@@

Model Overview

@@

•	 $P(Y=1 | X)$  represents the probability of Trump winning (i.e.,  Y=1 ).
•	 $\beta_0$  is the intercept term.
•	 $\beta_1$  and  $\beta_2$  are the coefficients for the predictor variables, which include:
•	 X_1 : The average polling percentage for Trump ( \text{trump\_pct} ).
•	 X_2 : The average polling percentage for Harris ( \text{harris\_pct} ).

Data and Features

@@

	•	trump_pct: The percentage of respondents supporting Trump.
	•	harris_pct: The percentage of respondents supporting Harris.
	•	sample_size: The size of each poll, which influences the weight of that poll in our analysis.
	•	start_date: The date when the poll began, which is used to calculate recency.

Weight Calculation

@@

Recency Weight:

@@

Total Weight:

@@

Logistic Regression Implementation

The logistic regression model is implemented using the LogisticRegression class from the sklearn library. The predictor variables are defined as:

@@

	•	Linearity: The log-odds of the outcome variable are linearly related to the predictor variables.
	•	Independence: Observations are independent of one another.

Limitations

	1.	Polling Bias: The model is contingent upon the quality of the polling data. Polls with lower numeric grades or transparency scores are excluded, potentially introducing bias if high-quality polls are not representative of the entire electorate.
	2.	Dynamic Voter Sentiment: Voter preferences can change rapidly, particularly in the lead-up to the election, and the model may not fully capture these shifts if not updated frequently.

Alternative Models Considered

Alternative models, such as Bayesian logistic regression, were considered. While Bayesian models allow for the incorporation of prior distributions and might provide additional insights through credible intervals, they require more complex implementation and careful selection of priors. Given the nature of the data and the objective of clear interpretability, the logistic regression model was chosen for its straightforward application and interpretability.

Conclusion

The logistic regression model provides a structured approach to predict the likelihood of Trump winning in various states, based on recent and high-quality polling data. By focusing on key features such as polling percentages and sample sizes, the model captures the competitive dynamics between candidates and offers valuable insights into potential election outcomes. Future iterations of this model can incorporate real-time data updates to enhance predictive accuracy as the election approaches.

@@

This approach, inspired by [@Stantcheva’s guide on survey creation](https://docs.google.com/forms/d/e/1FAIpQLSc5LYCAd0OmiIc-LHRFUVgVGuSMKzayVWZB7VOrsrDOe5742w/viewform?vc=0&c=0&w=1&flr=0), leverages financial incentives to align respondent predictions with their genuine expectations. By combining innovative sampling, incentivized engagement, and rigorous data validation, this methodology aims to bridge the gap between traditional polling and betting market predictions, ultimately enhancing the accuracy of our election forecast.

>>>>

The data used in this paper was gathered from FiveThirtyEight [@FiveThirtyEight] a website that aims to use "data and evidence to advance public knowledge". The programming language for data analysis, visualization and statistical investigation was Python [@Python] along with the packages Matplot [@Matplot], Seaborn [@Seaborn], Numpy [@Numpy], Pandas [@Pandas], Jupyter [@Jupyter], Tabulate [@Tabulate], Sklearn [@Sklearn]...

@@

# 2.2 The Dataset

A swing state in the context of a US election is defined as "a state where the number of Democratic and Republican voters is about the same, that has an important influence on the result of the election of the United States President". States that do not fall under his definition are not likely to "swing" to another party as we approach the election, and thus their outcome can be reasonably assumed. Thus the election result is normally decided by which party wins just a handful of states in the electoral college system. By concentrating on these swing states, we aim to capture the most critical and uncertain areas of the electoral map, where voter preferences are most likely to sway the final result. Thus our analysis restricts the datasets to state level polls in the following states: Arizona, Pennsylvania, North Carolina, Georgia, Nevada, Michigan, and Wisconsin [@nyt_swing, @FiveThirtyEight]. In figure 3 we see the ammount that each party hold according to those sources, with the remaining swing states, seen in the middle of this graph set to determine if either Harris or Trump recieve over 270th electoral votes.

@@

# 2.3 Variables of Interest
In our logistic regression model predicting the 2024 U.S. Presidential Election outcome, the primary variables of interest are derived from aggregated state-level polling data. The percentage of respondant favouring a candidate become the key variables for prediction. Within the collection of polls in our sample the independent variable `trump_pct` represents the percentage of respondents who indicate support for Donald Trump , and `harris_pct` respectively.
Polls ar taking towards the build up of an election and thus ar taken by repondants at different times. Graph  x shows how the sample of polls are distributed over time in which it is clear that we have a collection of polls from a wide range of periods. Because sentiment changes over time we will adjust the weight that we place on polls that are closer to the election as these will be more indicative of how respondants will be voting on the day.
```{python}
#| echo: false

@@

## Model Overview

@@

•	 $P(Y=1 | X)$  represents the probability of Trump winning (i.e.,  Y=1 ). \newline
•	 $\beta_0$  is the intercept term. \newline
\newline $\beta_1$  and  $\beta_2$  are the coefficients for the predictor variables, which include: \newline
•	 X_1 : The average polling percentage for Trump ( \text{trump\_pct} ). \newline
•	 X_2 : The average polling percentage for Harris ( \text{harris\_pct} ). \newline

## Data and Features

@@

•	trump_pct: The percentage of respondents supporting Trump. \newline
•	harris_pct: The percentage of respondents supporting Harris. \newline
•	sample_size: The size of each poll, which influences the weight of that poll in our analysis. \newline
•	start_date: The date when the poll began, which is used to calculate recency.

## Weight Calculation

@@

### Recency Weight:

@@

### Total Weight:

@@

### Logistic Regression Implementation

The logistic regression model is implemented using the LogisticRegression class from the sklearn library. The predictor variables are defined as: \newline

@@

•	Linearity: The log-odds of the outcome variable are linearly related to the predictor variables. \newline
•	Independence: Observations are independent of one another.

### Limitations

1.	Polling Bias: The model is contingent upon the quality of the polling data. Polls with lower numeric grades or transparency scores are excluded, potentially introducing bias if high-quality polls are not representative of the entire electorate. \newline 
2.	Dynamic Voter Sentiment: Voter preferences can change rapidly, particularly in the lead-up to the election, and the model may not fully capture these shifts if not updated frequently.

### Alternative Models Considered

Alternative models, such as Bayesian logistic regression, were considered. While Bayesian models allow for the incorporation of prior distributions and might provide additional insights through credible intervals, they require more complex implementation and careful selection of priors. Given the nature of the data and the objective of clear interpretability, the logistic regression model was chosen for its straightforward application and interpretability.\newline

Ultimately the logistic regression model provides a structured approach to predict the likelihood of Trump winning in various states, based on recent and high-quality polling data. By focusing on key features such as polling percentages and sample sizes, the model captures the competitive dynamics between candidates and offers valuable insights into potential election outcomes. Future iterations of this model can incorporate real-time data updates to enhance predictive accuracy as the election approaches.

@@

This approach, inspired by Stantcheva’s guide on survey creation [@Stantcheva], leverages financial incentives to align respondent predictions with their genuine expectations. By combining innovative sampling, incentivized engagement, and rigorous data validation, this methodology aims to bridge the gap between traditional polling and betting market predictions, ultimately enhancing the accuracy of our election forecast.
